% !TEX root = ../main.tex
\chapter{Introduction}

This document provides a standalone engineering report system for the ball and beam
experiment contained in this repository. The report combines the canonical modeling
record in \path{modeling.md}, calibration evidence from the April 7, 2026 beam-angle
campaign, and the staircase telemetry run stored in
\path{telemetry_logs/20260411_142117/telemetry.csv}.

The report structure is intentionally self-contained. All LaTeX sources, bibliography
files, generated figures, generated tables, and build outputs live inside
\path{report/}. The analysis pipeline reads the raw CSV logs directly and exports
report-ready plots and tabular summaries without depending on any other repository.

\section{Purpose}

The main goals are:

\begin{itemize}
  \item capture the first-principles ball and beam model used for the project,
  \item document the implemented sensing, calibration, and cascaded controller in
  the firmware,
  \item analyze the available calibration and staircase telemetry logs directly from
  raw CSV data, and
  \item provide a maintainable report build flow that works from the terminal and in
  VS Code with LaTeX Workshop.
\end{itemize}

\section{Data Sources}

The report uses the following repository-local sources as its primary experiment record:

\begin{itemize}
  \item \path{modeling.md} for the mathematical derivation, sign conventions, and
  implementation record,
  \item \path{src/main.cpp} for the shipped controller constants and mode logic,
  \item \path{telemetry_logs/20260411_142117/telemetry.csv} for the closed-loop
  staircase experiment, and
  \item \path{calibration_logs/combined_20260407_forward_reverse/combined_cleaned_calibration.csv}
  for beam-angle calibration hysteresis characterization.
\end{itemize}

\section{Report Roadmap}

Chapter~2 summarizes the physical apparatus, coordinates, and measured parameters.
Chapter~3 reproduces the nonlinear derivation from the canonical model record.
Chapter~4 specializes the model around the horizontal operating point for control
design. Chapter~5 documents the firmware implementation actually used in the hardware
run. Chapter~6 analyzes the beam-angle calibration data. Chapter~7 discusses the April
11 telemetry run using plots and performance metrics generated directly from the raw
CSV logs. Chapter~8 closes with key findings and immediate next steps.
